---
title: "Using a custom controlled vocabulary for rMZQC"
author: "Chris Bielow <chris.bielow@fu-berlin.de>"
date: '`r Sys.Date()`'
output:
  html_document: default
  pdf_document: null
vignette: >
  %\VignetteIndexEntry{Using a custom controlled vocabulary for rMZQC}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Using a custom controlled vocabulary for rMZQC

This vignette serves as a guide for R users to use a custom CV before creating an mzQC document.

Warning: you should settle on a CV before instantiating any mzQC objects, since this ensures that all
CV terms are consistent (and checked for existance) and that the CV meta information within the mzQC document is accurate.

**Target Audience:** R users


## Create a minial mzQC document with a custom CV

Let's first consider what happens by default:

```{r, eval=TRUE}
library(rmzqc)

print(getCVInfo())
## test if the default CV is usable
toQCMetric(id = "MS:4000059", value = 13405) ## number of MS1 scans

```

However, if you happen to run this code without an internet connection, it will fall back to
the PSI-MS CV which is shipped with the rmzqc package (which may not contain the latest CV terms)

```{r, eval=TRUE}
## With internet:
myCV = getCVSingleton()
myCV$setData(getCVDictionary("latest")) ## this is done internally by default when you load the package
cat("Number of entries in latest CV: ", nrow(getCVSingleton()$getCV()), "\n")
print(getCVInfo())

## simulate missing internet connection by invoking the function manually
myCV$setData(getCVDictionary("local"))
cat("Number of entries in local CV: ", nrow(getCVSingleton()$getCV()), "\n")
print(getCVInfo())
```

Now, the package's PSI-MS CV might still not suit you, and you want to use the latest unpublished CV, which you downloaded somewhere, or which you handcrafted for testing. Then simply use a custom .obo file:

```{r, eval=TRUE}
myOBO = system.file("./cv/psi-ms.obo", package="rmzqc") ## we will use a local file, but you can point to anything you have (even URI's)
myCV$setData(getCVDictionary("custom", myOBO))
cat("Number of entries in custom CV: ", nrow(getCVSingleton()$getCV()), "\n")
print(getCVInfo())

## you may want to change the CV-entries, or URI or version manually, before creating an mzQC file:
newCV = list(CV = myCV$getData()$CV, 
             URI = "https://myURI.com",
             version = "9.9.2")
myCV$setData(newCV)
print(getCVInfo())



```


